Distributional Trees and Forests

### 2017-10-10 pkg/disttree/inst/draft/draft.Rnw
changed tables and skills score plot
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1648 255f4249-2263-4417-b1ea-8498779e27a7

\author{Lisa Schlosser\\Universit\"at Innsbruck
   \And Torsten Hothorn\\Universit\"at Z\"urich
   \And Achim Zeileis\\Universit\"at Innsbruck
   \And Reto Stauffer\\Universit\"at Innsbruck}
\Plainauthor{Lisa Schlosser, Torsten Hothorn, Achim Zeileis, Reto Stauffer}

@@

\newpage
\begin{tabular}{| l  c  l |}
\hline
\thead{basic covariates} & \thead{nr} & \thead{variations}\\
\hline
\textbf{tp{\_}pow}: total precipitation    & 12 variations:  & ensemble mean of sum, \\
\hspace*{\fill} power transformed (by 1.6) &    & ensemble std dev of sum, \\ 
\textbf{cape{\_}pow}: convective           &    & ensemble min of sum, \\
\hspace*{0.4cm} available potential energy &    & ensemble max of sum\\
\hspace*{\fill}  power transformed (by 1.6)&    & \qquad all over 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sum over\\
\hspace*{\fill}                            &    & \qquad 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std dev of sum over\\
\hspace*{\fill}                            &    & \qquad 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\textbf{dswrf}: downwards short ware      & 6 variations: & ensemble mean of mean, \\
\hspace*{\fill} radiation flux (sunshine) &   & ensemble mean of min,\\
\textbf{msl}: mean sea level pressure     &   &  ensemble mean of max,\\
\textbf{pwat}: preciptable water          &   & ensemble std dev of mean,\\
\textbf{tmax}: 2m maximum                 &   & ensemble std dev of min,\\
\hspace*{1.15cm} temperature              &   & ensemble std dev of max,\\
\textbf{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{1.05cm} condensate  &   & \\
\textbf{t500}: temperature   &   & \\
\hspace*{1cm} on 500 hPA     &   & \\
\textbf{t700}: temperature   &   & \\
\hspace*{1cm} on 700 hPA     &   & \\
\textbf{t850}: temperature   &   & \\
\hspace{1cm} on 850 hPA      &   & \\
\hline
\textbf{tdiff500850}: temperature         & 3 variations: & ensemble mean of diff in mean,\\
\hspace*{\fill} difference 500 to 850 hPA &   & ensemble min of diff in mean,\\
\textbf{tdiff500700}: temperature         &   & ensemble max of diff in mean\\
\hspace*{\fill} difference 500 to 700 hPA &   & \qquad all over 6--30 UTC\\
\textbf{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPA &   & \\
\hline
\textbf{msl{\_}diff}: mean sea level  & 1 variation: & msl{\_}mean{\_}max - msl{\_}mean{\_}min\\
\hspace{1.65cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}

@@

\begin{tabular}{| l | c   c   c   c |}
\hline
 & \textbf{mu link} & \textbf{sigma link} & \textbf{mu covariates} & \textbf{sigma covariates} \\
\hline
\textbf{disttree} & identity & log & all covariates & all covariates \\ 
\hline
\textbf{distforest} & identity & log & all covariates & all covariates \\
\hline
\textbf{gamlss} & identity & log & tppow{\_}mean, & tppow{\_}sprd,\\
& & & tppow{\_}max, & dswrf{\_}sprd{\_}mean, \\
& & & tppow{\_}mean1218, & tppow{\_}sprd1218,\\ 
& & & capepow{\_}mean1218, & capepow{\_}mean1218,\\
& & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
& & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
& & & pwat{\_}mean{\_}mean, & \\
& & & tdiff500850{\_}mean, & \\
& & & msl{\_}diff & \\
\hline
\textbf{gamboostLSS} & identity & log & all covariates & all covariates \\
\hline
\textbf{EMOS} & identity & log & tppow{\_}mean & tppow{\_}spread \\
\hline
\end{tabular}
\\
For the gamlss model the covariates were selected manually based on expert knowledge. Additionally, interactions are specified in advance, once between tppow{\_}mean1218 and capepow{\_}mean1218 for the location parameter and once between tppow{\_}sprd1218 and capepow{\_}mean1218 for the l ocatioin parameter. By putting extra effort in the specification of the formulas the gamlss model is provided with additional information and is therefore expected to profit from this advantage.\\

@@

\begin{figure}[t!]
<<rain_pred_skills_score, fig=TRUE, echo=FALSE>>=
# skills score
  boxplot(1 - crps/crps[,6])
  abline(h = 0, col = "red")
@
\caption{\label{boxplots_rain}Results of predictions over all stations: CRPS skills score with reference model EMOS log}
\end{figure}

>>>>

\author{Lisa Schlosser\\Universit\"at Innsbruck
   \And Torsten Hothorn\\Universit\"at Z\"urich
   \And Reto Stauffer\\Universit\"at Innsbruck
   \And Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Lisa Schlosser, Torsten Hothorn, Reto Stauffer, Achim Zeileis}

@@

\newpage
\begin{table}[h]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{| l  c  l |}
\hline
Basic covariates & {\#} & Variations\\
\hline
\textbf{tp}: total precipitation            & 12 variations:  & ensemble mean of sum, \\
\hspace*{\fill} power transformed (by 1.6) &    & ensemble std dev of sum, \\ 
\textbf{cape}: convective                  &    & ensemble minimum of sum, \\
\hspace*{0.4cm} available potential energy &    & ensemble maximum of sum\\
\hspace*{\fill}  power transformed (by 1.6)&    & \qquad all over 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sum over\\
\hspace*{\fill}                            &    & \qquad 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std dev of sum over\\
\hspace*{\fill}                            &    & \qquad 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\textbf{dswrf}: downwards short ware      & 6 variations: & ensemble mean of mean values, \\
\hspace*{\fill} radiation flux (sunshine) &   & ensemble mean of minimal values,\\
\textbf{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\textbf{pwat}: preciptable water          &   & ensemble std dev of mean values,\\
\textbf{tmax}: 2m maximum                 &   & ensemble std dev of minimal values,\\
\hspace*{1.15cm} temperature              &   & ensemble std dev of maximal values,\\
\textbf{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{1.05cm} condensate  &   & \\
\textbf{t500}: temperature   &   & \\
\hspace*{1cm} on 500 hPA     &   & \\
\textbf{t700}: temperature   &   & \\
\hspace*{1cm} on 700 hPA     &   & \\
\textbf{t850}: temperature   &   & \\
\hspace{1cm} on 850 hPA      &   & \\
\hline
\textbf{tdiff500850}: temperature         & 3 variations: & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPA &   & ensemble minimum of difference in mean,\\
\textbf{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPA &   & \qquad all over 6--30 UTC\\
\textbf{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPA &   & \\
\hline
\textbf{msl{\_}diff}: mean sea level  & 1 variation: & msl{\_}mean{\_}max - msl{\_}mean{\_}min\\
\hspace{1.65cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}
\caption[Table caption text]{Basic covariates together with the number and the type of variations}
\label{covariates}
\end{minipage} 
%}
\end{table}

@@

\begin{table}[h]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{| l | l   l   l |}
\hline
 & \textbf{Type} & \textbf{Location ($\mu$)} & \textbf{Scale ($\sigma$)} \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{GAM (prespecified)} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\cdot$ & tp{\_}sprd1218 $\cdot$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{GAM (boosted)} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter}
\label{model_specification}
\end{minipage} 
%}
\end{table}

For the gamlss model the covariates were selected manually based on expert knowledge. Additionally, interactions are specified in advance, once between tppow{\_}mean1218 and capepow{\_}mean1218 for the location parameter and once between tppow{\_}sprd1218 and capepow{\_}mean1218 for the l ocatioin parameter. By putting extra effort in the specification of the formulas the gamlss model is provided with additional information and is therefore expected to profit from this advantage.\\

@@

# skills score
s <- 1 - crps[, 2:4]/crps[,6]
@

\begin{figure}[t!]
<<rain_pred_skills_score, fig=TRUE, echo=FALSE>>=
  matplot(t(s), type = "l", lwd = 1, col = gray(0.5, alpha = 0.5), lty = 1, axes = FALSE, xlab = "", ylab = "", xlim = c(0.5, 3.5))
  boxplot(s, add = TRUE)
  abline(h = 0, col = "gold", lwd = 2)
@
\caption{\label{boxplots_rain}Results of predictions over all stations: CRPS skills score with reference model EMOS log}
\end{figure}
